---
title: "Assignment 2 markdown"
output: html_document
---

```{r}
library(tidyverse)
```

```{r}
library(Stat2Data)
data("Hawks")
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
```
##1.1

```{r}
head(hawksSmall)
```
1. Discrete
2. Categorical
3. Categorical
4. Continuous
5. Continuous 

Issue with that plot is that it is trying to convey too much information with several visual cues, making it messy to read.

## 1.2
It also has no axes labels or title.

## 1.3
```{r}
hist <- ggplot(data=hawksSmall, aes(x=Weight)) + xlab("Weight")
hist+geom_histogram(binwidth=100)+ylab("Count")
```
       )
This univariate plot maps weight to horizontal position as the aesthetic.

It is bimodal as it has two peaks.

## 1.4
With adjust= 1:
```{r}
hist_tail<-ggplot(data=hawksSmall, aes(x=Tail))+xlab("Tail (mm)")
hist_tail + geom_density(adjust=1)+ylab("Density")
```
With adjust = 0.5:
```{r}
hist_tail + geom_density(adjust=0.5)+ylab("Density")
```
With adjust = 1 there are fewer peaks, so has lower mode (bimodal). With adjust = 0.5 it becomes trimodal.
```{r}
hist_tail_species <- ggplot(data=hawksSmall, aes(x=Tail, color=Species))
hist_tail_species+geom_density()+theme_bw()+xlab("Tail(mm)") + ylab("Density")
```

```{r}
ggplot(data=hawksSmall,aes(x=Tail,y=Species,fill=Species)) + geom_violin() + theme_bw() + xlab("Tail (mm)")
```
## 1.5

Aesthetics:
1. Position to tail
2. Position to weight
3. Colour to species
4. Shape to species

The glyphs are the basic shapes used to represent the species - triangle, square, circle

```{r}
ggplot(data=hawksSmall, aes(x=Tail, y=Weight, color=Species, shape=Species))+geom_point() + xlab("Tail (mm)") + ylab("Weight (g)")

```
## 1.6
```{r}
plot <- ggplot(data=hawksSmall, aes(x=Tail, y=Weight, color=Species)) + xlab("Tail (mm)") + ylab("Weight (g)") + geom_point() +  facet_wrap(vars(Species)) 

plot + geom_smooth(method="lm")
```
Visual cues here are colour and position.

In all the above plots, there appears to be a positive correlation between weights and tail lengths within hawks.

## 2.1
```{r}
hSF <- select(filter(hawksSmall, Species=="RT", Weight>=1000) ,Wing,Weight,Tail)
head(hSF)

```

```{r}
hawksSmall %>% 
  filter(Species=="RT", Weight>=1000) %>%
  select(Wing, Weight, Tail)
```
Has 3 variables, those which we chose with the select function. Could communicate by showing the head function or providing a legend table explaining each var.
Has 398 rows, so 398 observations and cases.

## 2.2

```{r}
head(arrange(hSF,Wing))
```
```{r}
species_name_full <- c("Cooper’s" , "Red-tailed",  "Sharp-shinned")
species_code <- c("CH", "RT", "SS")
spec_names_df <- data.frame(species_code,species_name_full)
head(spec_names_df)
```
```{r}
spec_names_df <- rename(spec_names_df, Species = species_code)
hawksFullName <- rename(select(left_join(Hawks, spec_names_df, by="Species"), -Species), Species=species_name_full)
head(select(hawksFullName, Species,Wing,Weight))
```
Type of join doesn't matter here as the relationship between the variables is one to one.
It would make a difference if there were many names for each code or many codes for each name, or worse, both.

## 2.4
```{r}
hawksWithBMI <- Hawks %>% 
  mutate(bird_BMI = 1000 * Weight/(Wing*Wing)) %>%
  select(Species, bird_BMI) %>%
  arrange(desc(bird_BMI))
head(hawksWithBMI)
```
```{r}

ggplot(filter(hawksWithBMI, bird_BMI <= 100), aes(x=bird_BMI, y=Species, fill=Species)) + xlab("Bird BMI") + ylab("Species") + theme_bw() + geom_violin()
```
## 2.5
```{r}
hawksFullName %>%
  group_by(Species)%>%
  summarise(
    num_rows=n(), mean_wing=mean(Wing,na.rm=TRUE), med_wing=median(Wing,na.rm=TRUE), t_mean_wing=mean(Wing, trim=0.1,na.rm=TRUE), tail_wing_ratio = mean(Wing/Tail,na.rm=TRUE))
```
```{r}
hawksFullName %>%
  group_by(Species)%>%
  select(Wing,Weight, Culmen,Hallux,Tail,StandardTail,Tarsus)%>%
  summarise(
    across(everything(), ~sum(is.na(.)))
  )

```
## 3.1

```{r}
Hawks %>%
  summarise(
    mean_wing = mean(Wing, na.rm=TRUE), t_mean_wing=mean(Wing,na.rm=TRUE,trim=0.1), wing_med= median(Wing, na.rm=TRUE), mean_weight=mean(Weight,na.rm=TRUE), t_mean_weight= mean(Weight,trim=0.1,na.rm=TRUE),Weight_med = median(Weight,na.rm=TRUE))

```
```{r}
Hawks %>%
  group_by(Species)%>%
  summarise(
    mean_wing = mean(Wing, na.rm=TRUE), t_mean_wing=mean(Wing,na.rm=TRUE,trim=0.1), wing_med= median(Wing, na.rm=TRUE), mean_weight=mean(Weight,na.rm=TRUE), t_mean_weight= mean(Weight,trim=0.1,na.rm=TRUE),Weight_med = median(Weight,na.rm=TRUE))

```
## 3.2


## 3.3

```{r}
hal <-Hawks$Hallux
hal<-hal[!is.na(hal)]
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)
mean(corrupted_hal)


```
```{r}
num_outliers_vect<-seq(0,1000)
means_vect<-c()
t_means_vect<-c()
medians_vect<-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  means_vect<-c(means_vect,mean(corrupted_hal))
  medians_vect<-c(medians_vect,median(corrupted_hal))
  t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim=0.1))
}


df_means_medians<-data.frame(num_outliers=num_outliers_vect,
mean=means_vect,t_mean=t_means_vect,
median=medians_vect)

df_means_medians%>%
pivot_longer(!num_outliers, names_to = "Estimator", values_to = "Value")%>%
ggplot(aes(x=num_outliers,color=Estimator,
linetype=Estimator,y=Value))+
geom_line()+xlab("Number of outliers")

```
```{r}
ggplot(Hawks, aes(x=Species, y=Weight)) + geom_boxplot()
```


```{r}
num_outliers <- function(sample){
  sum<-0
  sample = na.omit(sample)
  quantiles = quantile(sample, probs=c(0.25,0.75),na.rm=1, names=FALSE)
  q25<- quantiles[1]
  q75<-quantiles[2]
  IQR<- q75-q25
  for (i in sample){
    if(i < q25-1.5*IQR){
      sum=sum+1
    }
    else if (i > q75+1.5*IQR){
      sum=sum+1
    }
  }
  return(sum)
}


Hawks %>% 
  group_by(Species)%>%
  summarise(
    #mean_weight=mean(Weight,na.rm=TRUE)
    num_outliers_weight=num_outliers(Weight)
  )
```
