---
title: "Assignment 7"
author: "Joe Anderson"
date: "15/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1: Student’s t-confidence intervals

If the sample mean was higher then the width of the confidence interval would stay the same, as both the lower and upper limits are equally dependent on the mean.

If the sample standard deviation was higher then the width would increase.
If the sample size was higher then the width would decrease.

```{r}
library(tidyverse)
library(Stat2Data)
data("Hawks")

weights = pull(Hawks, Weight)
weights = na.omit(weights)
alpha<-0.01 # 99% confidence interval
n<-length(weights)
t<-qt(1-alpha/2,df=n-1)
l<-mean(weights)-t/sqrt(n)*sd(weights)
u<-mean(weights)+t/sqrt(n)*sd(weights)
c(l,u)
    
```
As this is a one-sample t-test we have assumed that the weights are distributed by a Gaussian.
```{r}
ggplot()+geom_density(data=data.frame(weights),aes(x=weights))

ggplot(data=data.frame(weights),aes(sample=weights))+theme_bw()+stat_qq()+stat_qq_line(colour="blue")
```
This appears to deviate from the Gaussian assumption, so maybe bootstrap is more appropriate.

## 2: One sample t-test
```{r}
library(palmerpenguins)

bill_adelie <- penguins%>%
  filter(species == "Adelie")%>%
  pull(bill_length_mm)

bill_adelie <- na.omit(bill_adelie)

# Significance level of 0.01
t.test(x=bill_adelie, mu=40, conf.level=0.99)
```
P value much smaller than 0.01, so reject null hypothesis and conclude that beak mean is not 40mm
Again assuming Gaussian and iid. 

##3: Implementing a one-sample t-test
```{r}

oneSampleT<-function(samples, mu){
  sample_mean <- mean(samples)
  sample_sd <- sd(samples)
  n <- length(samples)
  t <- (sample_mean - mu)/(sample_sd/sqrt(n))
  p <- 2*(1-pt(abs(t),df=n-1))
  return(p)
}

oneSampleT(bill_adelie,40)
```
##4: The paired t-test
```{r}
library(PairedData)
data("Barley")

t.test(x=pull(Barley,Glabron), y=pull(Barley,Velvet), paired=TRUE, conf.level=0.01)
```
p value greater than significance level, therefore accept null hypothesis, so true difference in mean is equal to 0.

```{r}
# Cohen's d
diffs <- pull(Barley,Glabron)-pull(Barley,Velvet)
s <- sd(diffs)
y_bar <- mean(diffs)
effect_size <- y_bar/s
effect_size
```

Assume that the differences are distribute normally for the paired case. Plot to see if justified:

```{r}
ggplot()+geom_density(data=data.frame(diffs),aes(x=diffs))

ggplot(data=data.frame(diffs),aes(sample=diffs))+theme_bw()+stat_qq()+stat_qq_line(colour="blue")
```

Appears to be a skewed Gaussian.

#5: Investigating coverage for Student's t intervals
```{r}
student_t_confidence_interval<-function(sample,confidence_level){
  sample<-sample[!is.na(sample)] # remove any missing values
  n<-length(sample) # compute sample size
  mu_est<-mean(sample) # compute sample mean
  sig_est<-sd(sample) # compute sample sd
  alpha = 1-confidence_level # alpha from gamma
  t<-qt(1-alpha/2,df=n-1) # get student t quantile
  l=mu_est-(t/sqrt(n))*sig_est # lower
  u=mu_est+(t/sqrt(n))*sig_est # upper
  return(c(l,u))
}

num_trials<-100000
sample_size<-30
mu_0<-1
sigma_0<-3
alpha<-0.05
set.seed(0) # set random seed for reproducibility
single_alpha_coverage_simulation_df<-data.frame(trial=seq(num_trials))%>%
mutate(sample=map(.x=trial,.f=~rnorm(n=sample_size,mean=mu_0,sd=sigma_0)))%>%
# generate random Gaussian samples
mutate(ci_interval=map(.x=sample,.f=~student_t_confidence_interval(.x,1-alpha)))%>%
# generate confidence intervals
mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0))))%>%
# check if interval covers mu_0
mutate(ci_length=map_dbl(.x=ci_interval, .f=~(max(.x)-min(.x))))
# compute interval length
single_alpha_coverage_simulation_df%>%
pull(cover)%>%
mean() # estimate of coverage probability
```
```{r}
num_trials<-100000
sample_size<-30
mu_0<-1
sigma_0<-3
alpha<-seq(0,1,0.01)
sample=rnorm(n=sample_size,mean=mu_0,sd=sigma_0)

alpha_coverage_simulation_df<-data.frame(alpha)%>%
# generate random Gaussian samples
mutate(ci_interval=map(.x=alpha,.f=~student_t_confidence_interval(sample,1-.x)))%>%
# generate confidence intervals
mutate(cover=map_lgl(.x=ci_interval,.f=~((min(.x)<=mu_0)&(max(.x)>=mu_0))))%>%
# check if interval covers mu_0
mutate(ci_length=map_dbl(.x=ci_interval, .f=~(max(.x)-min(.x))))
# compute interval length
alpha_coverage_simulation_df%>%
pull(cover)%>%
mean() # estimate of coverage probability

head(alpha_coverage_simulation_df)
ggplot(data=alpha_coverage_simulation_df, aes(x=1-alpha,y=cover)) + geom_line()
```
```{r}
ggplot(data=alpha_coverage_simulation_df, aes(x=1-alpha, y=ci_length)) + geom_line()
```

