---
title: "Assignment 5 Markdown"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

##1

$$p_X(x) = \{\alpha \text{ if } x = 1, \beta \text{ if } x = 5, 0 \text{ otherwise} \}$$

$$E(x) = \alpha + 5\beta$$

$$\text{Var}(x) = \alpha + 25\beta - \alpha^2 - 10\alpha\beta - 25\beta^2$$

##2

$$P(U\in[a,b]) = \int_a^b f_X(x)dx = b-a$$
```{r}


set.seed(0)

n<-1000

sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
    (0<=U)&(U<0.25)~1,
    (0.25<=U)&(U<0.5)~5,
    (0.5<=U)&(U<=1)~0))%>%
  pull(X)

```

Creates iid sequence using runif function, which selects n values from uniform distribution. These values are then mapped to the 1 and 5 probabilities using the inequalities below - first 25% map to 1, second 25% to 5, rest to 0.

```{r}
sample_X_015 = function(alpha,beta,n){
  sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
    (0<=U)&(U<alpha)~1,
    (alpha<=U)&(U<beta+alpha)~5,
    (beta+alpha<=U)&(U<=1)~0))%>%
  pull(X)
}

sample = sample_X_015(0.5,0.1,10000)
sample_avg = mean(sample)
sample_avg
```
$$E(x) = \alpha + 5\beta = 0.5 + 5\times 0.1 = 1$$
The simulation gets approximately the same expected value. The law of large numbers tells us that, as we increase n, the simulated average shall approach the expected value.

```{r}
sample_var = var(sample)
sample_var
```
$$\text{Var}(X) = \alpha + 25\beta - \alpha^2 - 10\alpha\beta - 25\beta^2 = 2$$
```{r}


sample_X_2 = function(beta){
  sample_X<-data.frame(U=runif(100))%>%
  mutate(X=case_when(
    (0<=U)&(U<0.1)~1,
    (0.1<=U)&(U<beta+0.1)~5,
    (beta+0.1<=U)&(U<=1)~0))%>%
  pull(X)
}

betas <- seq(0,0.9,0.01)
df1 <- data.frame(betas)%>%
  mutate(samples=map(betas,sample_X_2))%>%
  mutate(average=map_dbl(samples,mean))
  
ggplot(data=df1, aes(x=betas, y=average))+ geom_point()

```



## 3

$$f_{\mu,\sigma} = \frac{1}{\sigma\sqrt{2\pi}}\text{exp}(-\frac{1}{2}(\frac{x-\mu}{\sigma})^2)$$

```{r}
x = seq(-4,6,0.01)
pnorm1 = pnorm(x,mean=1,sd=sqrt(1))
pnorm2 = pnorm(x,mean=1,sd=sqrt(2))
pnorm3 = pnorm(x,mean=1,sd=sqrt(3))

qnorm_df = data.frame(x,pnorm1,pnorm2,norm3)%>%
  pivot_longer(!x,names_to="Variance",values_to="CDF")
ggplot(data=qnorm_df, aes(x=x, y=CDF, colour=Variance, linetype=Variance))+ geom_line()

```
See the relationship between the size of the tail and the rate of growth of the quantile function.
```{r}
x = seq(0,1,0.01)
qnorm1 = qnorm(x,mean=1,sd=sqrt(1))
qnorm2 = qnorm(x,mean=1,sd=sqrt(2))
qnorm3 = qnorm(x,mean=1,sd=sqrt(3))

norm_df = data.frame(x,qnorm1,qnorm2,qnorm3)%>%
  pivot_longer(!x,names_to="Variance",values_to="Density")
ggplot(data=norm_df, aes(x=x, y=Density, colour=Variance))+ geom_line()

```

```{r}
set.seed(0)
standardGaussianSample <- rnorm(100)
mean1Var3GaussianSampleA <- 1 + sqrt(3)*standardGaussianSample

 set.seed(0)
mean1Var3GaussianSampleB <- rnorm(100, 1, sqrt(3))

all.equal(mean1Var3GaussianSampleA,mean1Var3GaussianSampleB)
```

```{r}
x=seq(-5,7,0.1)
density = dnorm(x,mean=1,sd=sqrt(3))

normframe = data.frame(x,density)
sampleframe = data.frame(mean1Var3GaussianSampleA)

trueVarPlot <- ggplot()+geom_line(data=normframe, aes(x=x, y=density, colour="blue"))+geom_density(data=sampleframe,aes(x=mean1Var3GaussianSampleA, colour="red"))


trueVarPlot#+samplePlot
```

```{r}
samplePlot
```

##4  Binomial

```{r}
set.seed(0)
n=50
p=0.7
x=seq(1,50)
binom_df = data.frame(x,pmf=dbinom(x,n,p))
head(binom_df,3)
```
```{r}
x1=seq(0,50,0.01)
m=50*0.7
s=sqrt(50*0.7*(1-0.7))

gaussian_df=data.frame(x1,pdf=dnorm(x1,m,s))
head(norm_df,3)
```

```{r}
colors<-c("Gaussian pdf"="red", "Binomial pmf"="blue")
fill<-c("Gaussian pdf"="white", "Binomial pmf"="white")

ggplot()+labs(x="x",y="Probability")+theme_bw()+geom_line(data=gaussian_df,aes(x1,y=pdf,color="Gaussian pdf"),size=2)+
# create plot of Gaussian density
  geom_col(data=binom_df,aes(x=x,y=pmf,color="Binomial pmf",fill="Binomial pmf"))+
  scale_color_manual(name = "", values=colors)+
  scale_fill_manual(name = "", values=fill)+
  xlim(c(20,50))

```
# Exponential

```{r}
my_cdf_exp <- function(x,lambda){
  if(x<0){
    return(0)
  }else{
  return(1-exp(-lambda*x))
}}

lambda<-1/2
map_dbl(.x=seq(-1,4),.f=~my_cdf_exp(x=.x,lambda=lambda))
```

```{r}
test_inputs<-seq(-1,10,0.1)
my_cdf_output<-map_dbl(.x=test_inputs,.f=~my_cdf_exp(x=.x,lambda=lambda))
inbuilt_cdf_output<-map_dbl(.x=test_inputs,.f=~pexp(q=.x,rate=lambda))
all.equal(my_cdf_output,inbuilt_cdf_output)
```


```{r}
 my_quantile_exp <- function(p, lambda){
     return((-1/lambda)*log(1-p))
 }


test_inputs<-seq(0,1,0.01)
my_quant_output<-map_dbl(.x=test_inputs,.f=~my_quantile_exp(p=.x,lambda=lambda))
inbuilt_quant_output<-map_dbl(.x=test_inputs,.f=~qexp(p=.x,rate=lambda))
all.equal(my_quant_output,inbuilt_quant_output)
```

# 6 Poisson




